---
title: "extrapolate_S_mat"
author: "Simon Schulte"
date: "22 November 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# load packages
```{r}
library(xlsx)
library(data.table)
require(devtools)
install_github('simonpedrobolivar/masterthesis/MRIOtools')
require(MRIOtools)
library(parallel)
library(RColorBrewer)
require(gdata)
```


# setting the constant parameters
```{r parameters}
colors <- brewer.pal(12, "Set3") # colors for plots
n_ind <- 163 # number of industries
n_countries <- 49 # number of countries (40 + 1 RoW)
years_obs <- 1995:2011 # years for which IO table exists
n_years_obs <- length(years_obs)
years_est <- c(2020, 2030)
years <- c(years_obs, years_est)
n_FDcat <- 4 # final demand sectors (household, non-gov, gov, capital formation)
n_GHGs <- 4 # number of relevant GHGs (CO2, CH4, N2O, SF6)


stressors_vec = read.xls("/home/simon/Dokumente/Master/master_thesis/Calculation/MRIO_Config.xlsx", sheet = 3, header = T, skip = 1)$stressor
GHG_indices <- c(27 , 28  ,29 , 71  ,72 , 73  ,74 , 75  ,76 , 77  ,78 , 96  ,97 ,427 ,428 ,429 ,430 ,431 ,433 ,439 ,441, 442)
GHG_names <- as.character(stressors_vec[GHG_indices])
path4plots <- "/home/simon/Dokumente/Master/master_thesis/figures/"
path2data <- "~/Dokumente/Master/master_thesis/data/EB3/Scenario_4/"
```


# load data
```{r}
setwd(path2data)
files = list.files(pattern="S_")[1:length(years_obs)]
S_obs_list <- lapply(files, function(x) fread(x, sep = ",")) # read the files. List of data.tables. Each element of the list represents ONE year. 
setwd("~/Dokumente/Master/master_thesis/Results/MRIO_Results/Timeseries 1995 – 2015_2017_11_20__19_16_11")
files = list.files(pattern="Midpoint footprints, Germany")
S_names <- fread(files[1], sep = "\t", header = T, select = c(1,2), skip = 0)
country_names <- unique(S_names$Country)
dim(S_obs_list[[1]]) # check data
```

# check timeseries
```{r}
S_obs_list2check <- lapply(S_obs_list, function(x) as.data.table(t(x))) # transpose
S_obs_list2check <- lapply(S_obs_list2check, function(x) setnames(x, GHG_names)) # set colnames
S_obs_list2check <- lapply(S_obs_list2check, function(x) cbind(S_names, x)) # add column for industry and country
```


## trend analysis
```{r}
total_em_list <- sapply(S_obs_list2check, function(x) colSums(x[,-c(1,2)]))

barplot(total_em_list, legend.text = F, col = colors, ylim = c(0, 6e10), names.arg = years_obs)
legend("topright", legend = GHG_names, fill = colors, ncol = 2, cex = 0.5)

# same only weighted according to GWP100
total_gwp_mat <- apply(total_em_list, 2, function(x) ifelse(substr(rownames(total_em_list), 1, 3) == "CH4", x * 25, ifelse(substr(rownames(total_em_list), 1, 3) == "N20", x * 298, 
ifelse(substr(rownames(total_em_list), 1, 3) == "SF6", x * 22800, x)))) 

barplot(total_gwp_mat, legend.text = F, col = colors, ylim = c(0, 6e10), names.arg = years_obs)
legend("topright", legend = GHG_names, fill = colors, ncol = 2, cex = 0.5)
```
Was auffällt: 
- light increase from 2007 to 2011
- sharp decline from 1998 to 1999 (almost halbiert)
- co2 combustion air mit abstand der relevanteste emissionsquelle

## GER industries (co2 combustion air)
```{r}
S_obs_GER <- sapply(S_obs_list2check, function(x) x[Country == "Germany"]$`CO2 - combustion - air`)
colnames(S_obs_GER) <- years_obs
S_obs_GER_scaled <- t(scaleTo100(t(S_obs_GER), 1995))
S_obs_GER_scaled[is.nan(S_obs_GER_scaled)] <- 0
S_obs_GER_scaled[is.infinite(S_obs_GER_scaled)] <- NA

for(i in 1:163){ 
   plot(years_obs, S_obs_GER_scaled[i,], col = i, type = "l", las = 2, xaxt = "n")
  axis(1, years_obs, las = 2)
  legend("topright", legend = S_names[Country == "Germany"]$Industry[i], bg = "white")
  readline(prompt="Press [enter] to continue")
}

for(i in 1:163){ 
  plot(years_obs, S_obs_GER[i,], col = i, type = "l", las = 2, xaxt = "n")
  axis(1, years_obs, las = 2)
  legend("topright", legend = S_names[Country == "Germany"]$Industry[i], bg = "white")
  print(S_obs_GER[i,])
  readline(prompt="Press [enter] to continue")
}
```
was auffällt: 
- some industries have same form of timeseries, only on different level (probably due to desaggregation)
- some industries have weird declines/increases, especially the agricultural sector. (e.g. Cultivation of oil seeds)

## time-series for same industries
```{r}
S_obs_byIndustry <-  sapply(S_obs_list2check, function(x) x[Industry == "Manufacture of basic iron and steel and of ferro-alloys and first products thereof"]$`CO2 - combustion - air`)
S_obs_byIndustry_ts <- ts(t(S_obs_byIndustry), start = years_obs[1], end = years_obs[17], names = country_names)
var_coef_fun <- function(ts) sd(ts) / mean(ts)
var_coef <- apply(S_obs_byIndustry_ts, 2, var_coef_fun)


#min <- min(S_obs_byIndustry_ts[,which(var_coef < 0.3)])
min <- min(S_obs_byIndustry_ts[which(S_obs_byIndustry_ts != 0)])

max <- max(S_obs_byIndustry_ts[,which(var_coef < 0.3)])


#plot
plot(S_obs_byIndustry_ts[,1], ylim = c(0, max(S_obs_byIndustry_ts)))
abline(h = max, col = "grey", lty = 2)
for(i in 2:n_countries) {
  lines(S_obs_byIndustry_ts[,i], col = i)
  legend("topleft", legend = country_names[i], col = i)
  readline(prompt="Press [enter] to continue")
}
legend("topright", legend = country_names, col = 1:n_countries, pch = 0.7, ncol = 3)


# a first try to scale time series
scale_gts_fun <- function(gts){
  min <- min(gts)
  max <- max(gts)
  res <- (gts - min) / (max - min)
  attr(res, "scale:min") <- min
  attr(res, "scale:max") <- max
  return(res)
  #(ts2 - min(ts)) / max(ts - min(ts))
}
unscale_gts_fun <- function(gts, min, max){
  if("scale:min" %in% attributes(gts)$names) min <- attributes(gts)$'scale:min'
  if("scale:max" %in% attributes(gts)$names) max <- attributes(gts)$'scale:max'
  return(gts * (max - min) + min)
 }
gts_scaled <- scale_gts_fun(S_obs_byIndustry_ts)
max(gts_scaled)
attributes(gts_scaled)$'scale:max'

round(unscale_gts_fun(gts_scaled) - gts, 10) # scale and unscale works ok. dif < 1e-9

plot(gts_scaled[,1], ylim = c(0, max(gts_scaled)))
abline(h = max, col = "grey", lty = 2)
for(i in 2:n_countries) {
  lines(gts_scaled[,i], col = i)
  legend("topleft", legend = country_names[i], col = i)
  readline(prompt="Press [enter] to continue")
}

ts(ts)
(ts(ts))
fitmodel_fun <- function(ts, years.obs, years.est){
  fm <- glm(ts ~ years.obs + I(years.obs^2) + I(years.obs^3), family = gaussian(link = "logit"))
  preds_scaled <- predict(fm, newdata = data.frame("years.obs" = years.est))
  return(as.numeric((exp(preds_scaled)/(1+exp(preds_scaled)))))
  
}
gts_new_scaled <- list()
preds <- ts(matrix(ncol = n_countries, nrow = n_years_est), start = years_est[1], end = years_est[length(years_est)], frequency = 1/10, names = country_names)
for(i in 1:n_countries){
 preds[,i] <- fitmodel(gts_scaled[,i], years_obs, years_est)
}

plot(gts_scaled[,1], ylim = c(0, max(gts_scaled)), xlim = c(1995, 2030))
for(i in 2:n_countries) {
  lines(gts_scaled[,i], col = i)
  points(preds[,i], pch = 16, col = i)
  legend("topleft", legend = country_names[i], col = i)
  readline(prompt="Press [enter] to continue")
}

                                     
                              
preds_unscaled <- unscale_gts_fun(preds, 
                                  min = attributes(gts_scaled)$'scale:min', 
                                  max = attributes(gts_scaled)$'scale:max')

for(i in 1:n_countries) {
  plot(gts[,i], ylim = c(0, max(c(gts[,i], preds_unscaled[,i]))), xlim = c(1995, 2030), main = country_names[i])
  lines(preds_unscaled[,i])
  readline(prompt="Press [enter] to continue")
}

plot(log(preds_unscaled[,1]), ylim = c(0, 1.1 * log(attributes(gts_scaled)$'scale:max')), type = "n")
abline(h = log(attributes(gts_scaled)$'scale:max'), lty = 2)
abline(h = 0, lty = 2)
for(i in 1:n_countries){
  points(log(preds_unscaled[,i]), col = i)
}
```






# extrapolate S
```{r}
S_est_list <- extrapolate_matrix(S_obs_list, years.est = years_est, years.obs = years_obs, parallel = T, n.cores = 3)
S_list <- c(S_obs_list, S_est_list)
#S_list <- S_obs_list
names(S_list) <- years
```

# write new S matrices
```{r}
for(i in 1:length(years_est)){
  write.table(S_est_list[[i]], paste0(path2data, "S_", years_est[i], ".txt"), row.names = F, col.names = F, sep = "\t") 
}
```





# analyse resulting S matrices
 TODO

```{r}
S_list <- lapply(S_list, function(x) cbind(GHG_names, x))
S_dt <- rbindlist(S_list, idcol = T)
S_dt[1:10, 1:10]
setnames(S_dt, ".id", "Year")

names <- names(S_dt)[-c(1,2)]
S_byGHG <- S_dt[,sum(names), by = .(GHG_names, Year)]

S_byIndustry <- S_dt[,sum(Households, NPISH, Government, GFCF), by = .(Year, Industry)]

#fd_names <- names(S_dt)[-c(1,2,3)]
S_byFDcat_Households <- S_dt[,sum(Households), by = .(Year)]
S_byFDcat_NPISH <- S_dt[,sum(NPISH), by = .(Year)]
S_byFDcat_Government <- S_dt[,sum(Government), by = .(Year)]
S_byFDcat_GFCF <- S_dt[,sum(GFCF), by = .(Year)]
#S_byFDcat_Exports <- S_dt[,sum(`Exports: Total (fob)`), by = .(Year)]

S_byFDcat <- merge(merge(S_byFDcat_Households, S_byFDcat_NPISH, by = "Year"), merge(S_byFDcat_Government, S_byFDcat_GFCF, by = "Year"), by = "Year")
names(S_byFDcat) <- colnames
S_byFDcat_mat <- apply(as.matrix(S_byFDcat), c(1,2), as.numeric)
rownames(S_byFDcat_mat) <- S_byFDcat_mat[,1]
S_byFDcat_mat <- S_byFDcat_mat[,-1]
```

## plots
```{r}
pdf(paste0(path4plots, "S_byFDcat_wGDP.pdf"))
stacked_area_plot(S_byFDcat_mat/1E3, agg = F, agg.level = 7, colorvec = colors, yaxt.steps = 500, xaxt.steps = 5, legend.pos = "topleft", ylab = "Final Demand GER [Billion EUR]", ylim.adjust = 150, legend.ncol = 1)
lines(GDP_DEU)
lines(years_obs, GDP_DEU_without_netExport)
legend(x = 2000, y = 3200, legend = "GDP (source: World Bank)", lty = 1, lwd = 2, box.col = "white", bg = "white")
dev.off()
```


# validate forecasts
### unscaled
```{r}
models <- c("glm", "glm2", "glm3") # model types
for(i in 1:length(models)){
  model_name <- paste0("S_valid_scaled_", models[i])
  temp <- validate_forecast(S_obs_list, year.split = 2000, years.obs = years_obs, parallel = T, n.cores = 3, type = models[i], scale = F)
  save(temp, file = paste0("/home/simon/Dokumente/Master/master_thesis/scripts/R/data/", model_name, ".RData"))
  rm(temp)
  cat("", i)
}
```

### scaled
```{r}
models <- c("glm", "glm2", "glm3") # model types
for(i in 1:length(models)){
  model_name <- paste0("S_valid_", models[i])
  temp <- validate_forecast(S_obs_list, year.split = 2000, years.obs = years_obs, parallel = T, n.cores = 3, type = models[i], scale = T)
  save(temp, file = paste0("/home/simon/Dokumente/Master/master_thesis/scripts/R/data/", model_name, ".RData"))
  rm(temp)
}
```







#### glm1
```{r}
if(!exists("S_valid_scaled_glm1")){
  try(load("/home/simon/Dokumente/Master/master_thesis/scripts/R/data/S_validation_glm1.RData"))
  if(!exists("S_valid_glm1")){
    S_valid_scaled_glm1 <- validate_forecast(S_obs_list, year.split = 2000, years.obs = years_obs, parallel = T, n.cores = 3, type = "glm", scale = T)
    save(S_valid_scaled_glm1, file = "/home/simon/Dokumente/Master/master_thesis/scripts/R/data/S_validation_scaled_glm1.RData")
  }
}
rm(S_valid_scaled_glm1)
```

#### glm2
```{r}
if(!exists("S_valid_scaled_glm2")){
  try(load("/home/simon/Dokumente/Master/master_thesis/scripts/R/data/S_validation_glm2.RData"))
  if(!exists("S_valid_glm2")){
    S_valid_glm2 <- validate_forecast(S_obs_list, year.split = 2000, years.obs = years_obs, parallel = T, n.cores = 3, type = "glm2", scale = T)
    save(S_valid_glm2, file = "/home/simon/Dokumente/Master/master_thesis/scripts/R/data/S_validation_scaled_glm2.RData")
  }
}
rm(S_valid_glm2)
```

#### glm3
```{r}
if(!exists("S_valid_glm3")){
  try(load("/home/simon/Dokumente/Master/master_thesis/scripts/R/data/S_validation_glm3.RData"))
  if(!exists("S_valid_glm3")){
    S_valid_glm3 <- validate_forecast(S_obs_list, year.split = 2000, years.obs = years_obs, parallel = T, n.cores = 3, type = "glm3", scale = T)
    save(S_valid_glm3, file = "/home/simon/Dokumente/Master/master_thesis/scripts/R/data/S_validation_scaled_glm3.RData")
  }
}
rm(S_valid_glm3)
```


### unscaled
#### glm1
```{r}
if(!exists("S_valid_glm1")){
  try(load("/home/simon/Dokumente/Master/master_thesis/scripts/R/data/S_validation_glm1.RData"))
  if(!exists("S_valid_glm1")){
    S_valid_glm1 <- validate_forecast(S_obs_list, year.split = 2000, years.obs = years_obs, parallel = T, n.cores = 3, type = "glm", scale = F)
    save(S_valid_glm1, file = "/home/simon/Dokumente/Master/master_thesis/scripts/R/data/S_validation_glm1.RData")
  }
}
```

#### glm2
```{r}
if(!exists("S_valid_glm2")){
  try(load("/home/simon/Dokumente/Master/master_thesis/scripts/R/data/S_validation_glm2.RData"))
  if(!exists("S_valid_glm2")){
    S_valid_glm2 <- validate_forecast(S_obs_list, year.split = 2000, years.obs = years_obs, parallel = T, n.cores = 3, type = "glm2", scale = F)
    save(S_valid_glm2, file = "/home/simon/Dokumente/Master/master_thesis/scripts/R/data/S_validation_glm2.RData")
  }
}
```

#### glm3
```{r}
if(!exists("S_valid_glm3")){
  try(load("/home/simon/Dokumente/Master/master_thesis/scripts/R/data/S_validation_glm3.RData"))
  if(!exists("S_valid_glm3")){
    S_valid_glm3 <- validate_forecast(S_obs_list, year.split = 2000, years.obs = years_obs, parallel = T, n.cores = 3, type = "glm3", scale = F)
    save(S_valid_glm3, file = "/home/simon/Dokumente/Master/master_thesis/scripts/R/data/S_validation_glm3.RData")
  }
}
```





