---
title: "validate_forecasts"
author: "Simon Schulte"
date: "13 Januar 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# function to validate forecasts
```{r}
call_validate_forecast <- function(mat_obs_list, models, year.split, years.obs, parallel = F, n.cores, scale = F, path, overwrite = F, progress.bar = T){
  for(i in 1:length(models)){
    if(scale){
      model_name <- paste0(substr(deparse(substitute(mat_obs_list)), 1, 1), "_valid_scaled_", models[i])
    }else{
      model_name <- paste0(substr(deparse(substitute(mat_obs_list)), 1, 1), "_valid_", models[i])
    }
    if(!overwrite & paste0(model_name, ".RData") %in% list.files(path)){
      readline(prompt=paste0(model_name, ".RData already exists. Press [esc] to interupt. If you want to overwrite the file press [enter]"))
    }
    temp <- validate_forecast(mat_obs_list, year.split = year.split, years.obs = years.obs, parallel = parallel, n.cores = n.cores, type = models[i], scale = scale, progress.bar = progress.bar)
    save(temp, file = paste0(path, model_name, ".RData"))
    rm(temp)
    cat("", i)
  }
}
```

# call function
```{r}
path <- "/home/simon/Dokumente/Master/master_thesis/scripts/R/data/"
models <- "gam1"
call_validate_forecast(Y_obs_list, models = models, year.split = 2000, years.obs = years_obs, path = path, parallel = T, n.cores = 3, scale = T, progress.bar = T)
```



# check outcome
## load files
```{r}
Y_valid <- list()
temp.space <- new.env()
Y_valid[["data"]][["ets"]]  <- get(load(paste0(path, "Y_valid_scaled_ets.RData"), temp.space),
                                   temp.space)
Y_valid[["data"]][["glm1"]] <- get(load(paste0(path, "Y_valid_scaled_glm.RData"), temp.space), temp.space)
Y_valid[["data"]][["glm2"]] <- get(load(paste0(path, "Y_validation_scaled_glm2.RData"), temp.space),
                                   temp.space)
Y_valid[["data"]][["glm3"]] <- get(load(paste0(path, "Y_validation_scaled_glm3.RData"), temp.space),
                                   temp.space)
rm(temp.space)

Y_valid_list <- get(load(paste0(path, "Y_valid_glm1.RData")))
Y_valid_sc_list <- get(load(paste0(path, "Y_valid_scaled_glm.RData")))
Y_valid_scaled <- add_names2vali_Y(Y_valid_sc_list, Y_names, as.matrix(Y_obs_list$`2011`))
Y_valid_unscaled <- add_names2vali_Y(Y_valid_list, Y_names, as.matrix(Y_obs_list$`2011`))
cor_mat_scaled <- pearson(Y_valid_scaled, ranks = ranks, years.forecast = year2future, parallel = T, n.cores = 3)
cor_mat_unscaled <- pearson(Y_valid_unscaled, ranks = ranks, years.forecast = year2future, parallel = T, n.cores = 3)
plot_cv(cor_mat_scaled, year_split)
plot_cv(cor_mat_unscaled, year_split)
plot(1,1)
```
load(paste0(path, "Y_valid_scaled_glm1.RData"))
Y_valid_list_full <- add_names2vali_Y(Y_valid_list, Y_names, as.matrix(Y_obs_list$`2011`))
cor_mat_glm1 <- pearson(Y_valid_list_full, ranks, year2future, parallel = T, n.cores = 3)
plot_cv(cor_mat_glm1)

```{r}
# add columns with names and ranks
Y_valid[["data"]] <- lapply(Y_valid[["data"]], add_names2vali_Y, names = Y_names, Y.obs4ranks = as.matrix(Y_obs_list$`2011`)) 

```


### by (1) and (2) (year where data is split and number of years to which forecasts are made)

```{r}
year <- 2000:2010
year2future <- 1:(length(year) - 1)
cor_mat1 <- matrix(nrow = length(year), ncol = 10)
colnames(cor_mat1) <- year2future
rownames(cor_mat1) <- year
 for(i in 1:length(year)){
   for(j in 1:length(year2future)){
     cor_mat1[i,j] <- cor(Y_valid_sc_list_full[Year_forecast - Year_split == year2future[j] & Year_split == year[i]]$est,  Y_valid_sc_list_full[Year_forecast - Year_split == year2future[j] & Year_split == year[i]]$obs)^2
   }
 }

par(mfrow = c(3,4))
 for(i in 1:nrow(cor_mat1)){
   plot(year2future, cor_mat1[i,], type = "b", ylim = c(0, max(cor_mat1, na.rm = T)), main = rownames(cor_mat1)[i])
 }
```

### by (2) and (3)
```{r}
ranks <- c(50, 100, 200, 7987)
 cor_mat2 <- matrix(nrow = length(ranks), ncol = 10)
 
  for(i in 1:length(ranks)){
   for(j in 1:length(year2future)){
     cor_mat2[i,j] <- cor(Y_valid_sc_list_full[Year_forecast - Year_split == year2future[j] & Rank < ranks[i]]$est,  Y_valid_sc_list_full[Year_forecast - Year_split == year2future[j]  & Rank < ranks[i]]$obs)^2
   }
 }
 par(mfrow = c(2,2))
 for(i in 1:nrow(cor_mat2)){
   plot(year2future, cor_mat2[i,], type = "b", ylim = c(0, max(cor_mat2, na.rm = T)), main = ranks[i])
 }
```


### by (1), (2) and (3)
all in one: 
```{r}
ranks  <- c(50, 100, 200, 7987)
year_split <- 2000:2010
year2future <- 1:(length(year) - 1)

# leaving time series with observation purely 0 in 
cl <- makeCluster(3)
clusterCall(cl, function() library(data.table))
Y_valid[["cor.arrays"]] <- parLapply(cl, Y_valid$data, pearson, ranks = ranks, years.forecast = year2future)
stopCluster(cl)

# excluding observation of 0
cl <- makeCluster(3)
clusterCall(cl, function() library(data.table))
Y_valid[["cor.arrays_w/o_zero"]] <- parLapply(cl, lapply(Y_valid$data, function(x) x[Rank != max(Rank)]), pearson, ranks = ranks, years.forecast = year2future)
stopCluster(cl)

plot_cv(Y_valid$cor.arrays$ets, year_split)
plot_cv(Y_valid$cor.arrays$glm1, year_split)


plot_cv(Y_valid$cor.arrays, year_split)

plot_cv(Y_valid$`cor.arrays_w/o_zero`, year_split)

 # excluding observations of 0 ?? (with dt[obs > 0]) ==> makes no difference
```

```{r}
par(mfrow = c(1,1))
plot(Y_valid_ets_full[Year_forecast - Year_split == 8 & Year_split == 2001 & Rank < 51]$est,  Y_valid_ets_full[Year_forecast - Year_split == 8  & Year_split == 2001 & Rank < 51]$obs, ylim = c(-4, 4), xlim = c(-4, 4))
cor(Y_valid_ets_full[Year_forecast - Year_split == 8 & Year_split == 2001 & Rank < 51]$est,  Y_valid_ets_full[Year_forecast - Year_split == 8 & Year_split == 2001 & Rank < 51]$obs)^2
lines(c(-4, 4), c(-4, 4))
```






